An Image-based method for phase estimation, gating and temporal super-resolution of \\cardiac ultrasound

### 2017-02-21 paper155.tex
Updated paper and figures -- finished kernel regression section

\begin{abstract}
In this paper, we present a novel image-based method for the estimation of instantaneous cardiac and respiratory phases from cardiac ultrasound videos and demonstrate its utility in gating and temporal super-resolution. We first use a novel strategy based on inter-frame similarity to transform the complex high-dimensional periodic cardiac ultrasound image sequence into a 1D time series with the same periodicity characteristics.  Next, we use a trend extraction technique to decompose this time series into a  trend component encoding respiratory motion alone and a residual component encoding the beating heart motion along. Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame from these signals. Next, we present a robust non-parametric regression technique that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion. Lastly, we use a kernel regression with a novel kernel to learn an image manifold parameterized by instantaneous cardiac phase that can be sampled to generate a single-cycle video at a higher temporal-resolution. We demonstrate and validate our methods using cardiac ultrasound videos and associated ECG recordings of 6 mice.
%\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal super-resolution}
\end{abstract}

@@

%\vspace{-0.5cm}
\section{Method}
\label{sec:method}
%
In this section, we present the theory underlying the proposed methods for instantaneous phase estimation, gating and temporal super-resolution of cardiac ultrasound videos. Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}. We then present a robust method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion in Section \ref{sec:method:gating}. Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression method that uses the phase-tagged image sequence to reconstruct a single-period video at higher-temporal resolution. 
%
%
\begin{IonFigT}
\centering
%
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
\ionbox{5.0in}\\
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}
\ionbox{5.0in}\\
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
\ionbox{5.0in}\\
%
\caption{Illustration of correspondence between ECG and the instantaneous cardiac phase estimated using the proposed method: (a) ECG signal (blue) and the cardiac phase estimated using our method using only image data. Also shown are the peaks of the QRS complex (pink circle) and the minima of the estimated cardiac phase (black circle) that correspond to the QRS peaks, (b) Five video frames evenly spaced in time between the first and second QRS peaks of the ECG signal constituting one cardiac cycle, and (c) Five video frames evenly spaced in time between first and second minima of the estimated cardiac phase constituting one cardiac cycle.}
\label{fig:instaphase_vs_ecg}
\end{IonFigT}
%\vspace{-0.5cm}
\subsection{Estimation of instantaneous cardiac and respiratory phases}
\label{sec:method:phase_estimation}
%
\begin{IonFigT}
\centering
%
\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
\ionbox{1.6in}\ionbox{1.6in}\ionbox{1.6in}\\
\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
\ionbox{4.75in}\\
\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/periodogram.png}
\ionbox{4.75in}\\
\includegraphics[width=4.75in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
\ionbox{4.75in}\\
%
\caption{Illustration of the proposed phase estimation method: (a) Inter-frame similarity matrix, (b) Trend component corresponding to the respiratory motion, (c) Seasonal component corresponding to the beating heart motion, (d) The similarity profile chosen for phase estimation along with the associated trend/respiration and seasonal/heart-beat components, (e) Periodogram of the heart-beat and respiration components along with the periodicity characteristics such as frequency and cycle duration, and (f) Instantaneous phase of the heart-beat and respiration components.}
\label{fig:phase_estimation}
\end{IonFigT}
%
While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data (wherein a single-variable is measured at each time point)\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting which is the case with periodic cardiac ultrasound videos wherein several thousands of variables (intensities of pixels) could be measured at each time point. Our strategy was to find a way to simplify this seemingly complex multi-variate problem by transforming it into a univariate one and take advantage of existing univariate methods to solve the problem at hand.

	We first compute the similarity between all pairs of frames in the given periodic image sequence containing $N$ images/frames to create a matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence. Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms \cite{Goshtasby2012} can potentially be chosen. Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames. If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted we can expect this time series to encompass the predominant periodicity characteristics of the original image sequence. Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion. Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos computed using the normalized correlation metric wherein these two sources of periodicity can be clearly observed in the frame similarity signal of many rows/columns.

	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Hodrick1997}, widely used in econometrics, to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion. The HP filter performs the decomposition of $u(t) = \tau_{resp}(t) + r_{heart}(t)$ by solving the following optimization problem:
\begin{equation}	
\argmin{\tau(t)} \left[ \sum_{t=1}^{N}  \left(u(t) - \tau_{resp}(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau_{resp}(t) \right)^2 \right] 
\end{equation}
where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative that penalizes curvature of the trend signal. We use a ridge-regression implementation of the HP filter~\cite{Yamada2015} with $\lambda=6400$. Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components, respectively, of the frame similarity signals in the corresponding rows of the similarity matrix $A$. Figures~\ref{fig:phase_estimation}(b,c) show the trend $A_{res[}$ and residual $A_{heart}$ matrices for one of our cardiac ultrasound videos. To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_{heart}$ whose periodogram or power-frequency distribution has minimum entropy. Let $\hat{u}(t)$ be the chosen frame similarity signal and let $\hat{\tau}_{resp}(t)$ and $\hat{r}_{heart}(t)$  be its associated trend/respiration and residual/heart-beat component signals, respectively. Figure~\ref{fig:phase_estimation}(d) shows the selected frame similarity signal (blue) along with the associated trend/respiration (pink) and residual/heart-beat (green) components. Figure~\ref{fig:phase_estimation}(e) shows the periodogram of these two components. Through Figures~\ref{fig:phase_estimation}(b-e), notice how well the two periodic sources have been decoupled from the raw frame similarity signal. Also, in Figure~\ref{fig:phase_estimation}, notice that most of the energy in the periodogram or power-frequency distribution of the respiration and heart-beat components is concentrated within a narrow frequency band. Furthermore, notice how these 1D signals capture the periodicity characteristics (e.g. frequency in beats-per-minute (bpm), period or cycle duration in frames, number of cycles) of the respective sources of motion in the input high-dimensional periodic image sequence.

	Next, considering the narrow-band nature of the derived trend/respiratory and residual/heart-beat signals, we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame. Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows:
\begin{equation}
\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)
\end{equation} 
and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$. Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases computed from the trend/respiration $\hat{\tau}_{resp}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signals, respectively, using the aforementioned approach. Figure~\ref{fig:phase_estimation}(f) shows the instantaneous cardiac and respiratory phases computed from the heart-beat and respiration signals in in Figure~\ref{fig:phase_estimation}(d).

%\vspace{-0.5cm}
\subsection{Gating out respiratory frames}
\label{sec:method:gating}
%
Once the instantaneous cardiac and respiratory phases of each frame has been estimated, it can be used to facilitate the extraction of quantitative measurements from a desired part/point of the periodic cycle, a process commonly referred to as gating. In this section, we present a robust two-step method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.

@@

	In the second step, we learn a robust mapping $S(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the residual/heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOESS) to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signal values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above. Next, we compute the median absolution value $\gamma_{MAD}$ of the difference between the LOESS fit $S( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames as a robust estimate of the dispersion of non-respiratory frames around the LOESS fit. Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert S( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \gamma_{mad}  \right \}$ for which the difference between the LOESS fit and heart-beat signal value is greater than $k \times \gamma_{mad}$ (we set $k = 2.0$).	
%
\begin{IonFigT}
\centering
%
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/respiratory_phase_gating.png}
\ionbox{5.0in}\\
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/robust_lowess_gating.png}
\ionbox{5.0in}\\
%
\caption{Illustration of our method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the trend/respiration, residual/heart-beat, and instantaneous respiratory phase signals (b) }
\label{fig:respiratory_gating}
\end{IonFigT}
%
%\vspace{-0.5cm}
\subsection{Learning to reconstruct image at any cardiac phase}
\label{sec:method:super_resolution}
%
Once the instantaneous cardiac phase of each frame has been estimated, it can be used to enable spatio-temporal super-resolution by considering the set of images within each period of a periodic image sequence as distinct observations of the same data. In this section, we present a method that uses a phase-tagged periodic image sequence to reconstruct a single-period video at a higher temporal resolution.

Given a de-noised periodic sequence $\{L_1, ..., L_n\}$ of $n$ images (with $m$ pixels each) along with their estimated intra-period phases $\{\phi_1, ..., \phi_n\}$, we first use Nadarya-Watson kernel regression\cite{Bishop2006} to construct a phase-parameterized image manifold in the form of a function $M(\phi): [0, 1) \to R^m $ that maps any intra-period phase $\phi$ to an image as follows:
\begin{equation}
M(\phi) = \frac{\sum_{i = 1}^{n} K(\phi, \phi_i) L_i}{\sum_{i = 1}^{n} K(\phi, \phi_i)} 
\end{equation}
where in we define the kernel $K$ using a gaussian with standard-deviation $\sigma$ equal to a multiple of the average distance between consecutive phases after sorting them. Note that, we account for the periodic nature of the intra-period phase whenever we measure distance between two phase values. A single-period video can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution from the range $[0, 1)$ of intra-period phase. Note that the aformentioned process will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in intra-period phase by some non-zero amount less than its temporal sampling interval. Also, the amount of temporal super-resolution possible will depend both on the number of periods observed and the maximum amount by which one can expect the images to be perturbed in phase across observations. 
%\vspace{-0.5cm}
\section{Results}
\label{sec:results}
%
%
\begin{IonFigT}
\centering
%
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
\ionbox{5.0in}\\
%
\caption{Visualization of m-mode frames ordered by estimated cardiac phase}
\label{fig:phase_ordering}
\end{IonFigT}
%
%
%\vspace{-0.5cm}
\section{Discussion and future work}
\label{sec:discussion}
%

>>>>

\begin{abstract}
In this paper, we present a novel purely image-based method for the estimation of instantaneous cardiac and respiratory phases from cardiac ultrasound videos and demonstrate its utility in gating and temporal super-resolution. We first use a novel strategy based on inter-frame similarity to transform the complex high-dimensional image sequence into a 1D time series with the same periodicity characteristics.  Next, we use a trend extraction technique to decompose this time series into a trend component encoding respiratory motion alone and a residual component encoding the beating heart motion alone. Next, we use the Hilbert transform to estimate the instantaneous cardiac and respiratory phases of each frame from these components. Next, we present a robust non-parametric regression technique that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion. Lastly, we use a kernel regression with a novel kernel to learn an image manifold parameterized by instantaneous cardiac phase that can be sampled to generate a single-cycle video at a higher temporal-resolution. We demonstrate and validate our methods using cardiac ultrasound videos and associated ECG recordings of 6 mice.
%\keywords{Cardiac, Ultrasound, Phase estimation, Gating, Temporal super-resolution}
\end{abstract}

@@

%\vspace{-0.5cm}
\section{Method}
\label{sec:method}
%
In this section, we present the theory underlying the proposed methods for instantaneous phase estimation, gating and temporal super-resolution of cardiac ultrasound videos. Specifically, we begin by describing our method for the estimation of instantaneous cardiac and respiratory phases in Section \ref{sec:method:phase_estimation}. We then present a robust method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion in Section \ref{sec:method:gating}. Finally, in Section \ref{sec:method:super_resolution}, we present a kernel regression method that uses the phase-tagged image sequence to reconstruct a single-period video at higher-temporal resolution. 
%
%
\begin{IonFigT}
\centering
%
\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/ecg_instaphase_overlay.png}\\
\ionbox{4.5in}\\
\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/qrs_peak_to_peak.png}\\
\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instaphase_valley_to_valley.png}
\ionbox{4.5in}\\
%
\caption{Illustration of correspondence between ECG and the instantaneous cardiac phase estimated using the proposed method: (a) ECG signal (blue) and the cardiac phase estimated using our method overlaid with peaks of the QRS complex (red circle) and the corresponding cardiac phase minima (black circle), (b) Five video frames evenly spaced in time between the first and second QRS peaks of the ECG signal (top-row) and the corresponding cardiac phase minima (bottom-row) constituting one cardiac cycle.}
\label{fig:instaphase_vs_ecg}
\end{IonFigT}
%\vspace{-0.5cm}
\subsection{Estimation of instantaneous cardiac and respiratory phases}
\label{sec:method:phase_estimation}
%
\begin{IonFigT}
\centering
%
\includegraphics[width=4.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/simMat.png}
\ionbox{1.33in}\ionbox{1.33in}\ionbox{1.33in}\\
\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/season_trend_decomposition.png}
\ionbox{4.5in}\\
\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/periodogram.png}
\ionbox{4.5in}\\
\includegraphics[width=4.5in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/instantaneous_phase.png}
\ionbox{4.5in}\\
%
\caption{Illustration of our phase estimation method: (a) Inter-frame similarity matrix, (b) Trend component matrix corresponding to respiratory motion, (c) Residual component matrix corresponding to beating heart motion, (d) Similarity profile chosen for phase estimation with associated trend/respiration and residual/heart-beat components, (e) Periodogram of heart-beat and respiration components along with periodicity characteristics (e.g. frequency, cycle duration) calculated for the peak frequency, and (f) Instantaneous cardiac and respiratory phases estimated using the Hilbert transform.}
\label{fig:phase_estimation}
\end{IonFigT}
%
While there have been numerous efforts for the estimation of instantaneous phase and/or frequency in periodic univariate time series data (wherein a single-variable is measured at each time point)\cite{Boashash1992,Lu2013,Luo2003}, there are not many methods that tackle this problem in a multivariate setting which is the case with periodic cardiac ultrasound videos wherein several thousands of variables (pixel intensities) could be measured at each time point. Our strategy was to find a way to simplify this complex multi-variate problem by transforming it into a univariate one and take advantage of existing univariate methods to solve the problem.

	We first compute the similarity between all pairs of frames in the given periodic image sequence containing $N$ images/frames to create a matrix $A \in R^{N \times N}$ where in the element $A(i,j)$ is equal to the similarity between the $i^{th}$ and $j^{th}$ frames of the given sequence. Here in, we use normalization correlation also known as the Pearson correlation coefficient to quantify inter-frame similarity but, in principle, any of the similarity metrics used in image registration algorithms \cite{Goshtasby2012} can potentially be chosen. Each row in the inter-frame similarity matrix $A$ can now be seen as a univariate time series that measures the similarity of the corresponding frame in the image sequence with all other frames. If the similarity metric is chosen with care and if the corresponding frame is not significantly corrupted we can expect this time series to encompass the predominant periodicity characteristics of the original image sequence. Two kinds of periodic motion are present in cardiac ultrasound videos, one corresponding to the beating heart motion and the other corresponding to respiratory motion. Figure \ref{fig:phase_estimation}(a) shows the inter-frame similarity matrix of one of our cardiac ultrasound videos computed using the normalized correlation metric wherein these two sources of periodicity can be clearly observed in many rows/columns.

	Next, we use a trend extraction technique~\cite{Alexandrov2012} called the Hodrick-Prescott (HP) filter~\cite{Hodrick1997}, widely used in econometrics, to decompose the frame similarity signal $u(t)$ corresponding to each row of $A$ into a sum of two components: (i) Trend component $\tau_{resp}(t)$ encompassing the periodicity characteristic of only the lower-frequency respiratory motion, and (ii) Residual component $r_{heart}(t)$ encompassing the periodicity characteristic of only the higher-frequency beating heart motion. The HP filter performs the decomposition of $u(t) = \tau_{resp}(t) + r_{heart}(t)$ by solving the following optimization problem:
\begin{equation}	
\argmin{\tau(t)} \left[ \sum_{t=1}^{N}  \left(u(t) - \tau_{resp}(t) \right)^2  + \lambda \sum_{t=1}^{N-1} \left( \nabla^2 \tau_{resp}(t) \right)^2 \right] 
\end{equation}
where $\nabla^2\tau_{resp}(t) = \tau_{resp}(t+1) - 2 \tau_{resp}(t) + \tau_{resp}(t-1)$ is the second-order difference or derivative that penalizes curvature of the trend signal. We use a ridge-regression implementation of the HP filter~\cite{Yamada2015} with $\lambda=6400$. Let $A_{resp}$ and $A_{heart}$ be the matrices whose rows contain the trend/respiratory and residual/heart-beat components, respectively, of the frame similarity signals in the corresponding rows of the similarity matrix $A$. Figures~\ref{fig:phase_estimation}(b,c) show the trend $A_{res[}$ and residual $A_{heart}$ matrices for one of our cardiac ultrasound videos. To minimize the effect of any noise on phase estimation, we pick frame similarity signal corresponding to the row of $A_{heart}$ whose periodogram or power-frequency distribution has minimum entropy. Let $\hat{u}(t)$ be the chosen frame similarity signal and let $\hat{\tau}_{resp}(t)$ and $\hat{r}_{heart}(t)$  be its associated trend/respiration and residual/heart-beat component signals, respectively. Figure~\ref{fig:phase_estimation}(d) shows the selected frame similarity signal (blue) along with the associated trend/respiration (pink) and residual/heart-beat (green) components. Figure~\ref{fig:phase_estimation}(e) shows the periodogram of these two components that are nicely segregated in frequency domain. The periodicity characteristics (e.g. frequency in beats-per-minute (bpm), cycle duration in frames, number of cycles) of heart-beat and respiration can be calculated based on the maximum-power frequency of their periodograms.

	Next, considering the narrow-band nature (see  Figure~\ref{fig:phase_estimation}) of the derived trend/respiratory and residual/heart-beat signals, we use the Hilbert transform~\cite{Lu2013} to estimate the instantaneous phase of each frame. Specifically, we compute the instantaneous intra-period phase $\phi(t) \in \left [  -\pi, \pi\right )$ of the periodic time series $x(t)$ using its Hilbert transform $H_x(t)$ as follows: $\phi(t) = arctan \left( \frac{H_x(t)}{x(t)}\right)$ and map $\phi(t)$ from the range $\left [  -\pi, \pi\right )$ to $\left [  0, 1\right )$. Let $\hat{\phi}_{heart}(t)$ and $\hat{\phi}_{resp}(t)$ denote the instantaneous cardiac and respiratory phases computed from the trend/respiration $\hat{\tau}_{resp}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signals, respectively. Figure~\ref{fig:phase_estimation}(f) shows the instantaneous cardiac and respiratory phases computed from the heart-beat and respiration signals in Figure~\ref{fig:phase_estimation}(d). 
%
%\vspace{-0.5cm}
\subsection{Gating out respiratory frames}
\label{sec:method:gating}
%
%
\begin{IonFigT}
\centering
%
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/respiratory_phase_gating.png}
\ionbox{5.0in}\\
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/robust_lowess_gating.png}
\ionbox{5.0in}\\
%
\caption{Illustration of our two-step method for gating out respiratory frames: (a) Frames $F_{cutoff}$ discarded (blue circles) in the first step overlaid with the trend/respiration, residual/heart-beat, and instantaneous respiratory phase signals (b) residual/heart-beat signal $\hat{r}_{heart}(t)$ vs cardiac phase $\hat{\phi}_{heart}(t)$ overlaid with frames $F_{cutoff}$ discarded in step-1 (blue circle), LOWESS fit $L(\phi)$ (black solid), upper and lower bounds or 95\% confidence interval $(L(\phi) \pm 2.0 * \hat{\sigma}_L)$ of valid non-respiratory frames (black dotted), and frames $F_{resp}$ (red circles) gated at the end of step-2.}
\label{fig:respiratory_gating}
\end{IonFigT}
%
Once the cardiac and respiratory phases of each frame has been estimated, it can be used to guide the extraction of quantitative measurements to a desired part/point of the periodic cycle, a process commonly referred to as gating. In this section, we present a robust two-step method that uses these phase estimates to gate out video frames predominantly influenced by respiratory motion.

@@

	In the second step, we learn a robust mapping $L(\phi) : \left [  0, 1\right ) \to R$ from cardiac phase to the residual/heart-beat component of the frame similarity signal by fitting a robust non-parametric regression model called Locally weighted regression (LOWESS) to the dataset $\left \{ \left(\hat{\phi}_{heart}(t), \hat{r}_{heart}(t) \right) \mid t \notin F_{cutoff}  \right \}$ containing the pair of the cardiac phase $\hat{\phi}_{heart}(t)$ and residual/heart-beat $\hat{r}_{heart}(t)$ signal values of all frames that do not belong to the set of frames $F_{cutoff}$ discarded in the first step above. Next, we compute a robust estimate of the standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit based on the median absolute value of the difference between the LOWESS fit $L( \hat{\phi}_{heart}(t) )$ and residual/heart-beat signal $\hat{r}_{heart}(t)$ for all frames. Lastly, we gate out all frames $F_{resp} = \left \{ t \; \big\lvert \; \lvert L( \hat{\phi}_{heart}(t) ) - \hat{r}_{heart}(t)  \rvert   > k \times \hat{\sigma}_{L}  \right \}$ for which the absolute  difference between the LOWESS fit and heart-beat signal value is greater than $k = 0.2$ times the standard deviation $\hat{\sigma}_{L}$.	
%
%\vspace{-0.5cm}
\subsection{Model to generate images by cardiac phase for super-resolution}
\label{sec:method:super_resolution}
%
In this section, we present a kernel regression model to reconstruct the image at any cardiac phase. This model can then be used to generate a single-cycle video representative of the subject's heart-beat at a desired higher temporal resolution to achieve temporal super-resolution.

Given a cardiac ultrasound video $I(t) : \{1, ..., N\} \to R^m$ of $N$ images with $m$ pixels each, we extract the residual/heart-beat component $\hat{r}_{heart}(t)$ of the frame similarity signal, estimate the instantaneous cardiac phase $\hat{\phi}_{heart}(t)$ of each frame, and compute the robust LOWESS fit $L(\phi) : \left [  0, 1\right ) \to R$ that maps cardiac phase to the heart-beat signal as  described in Sections~\ref{sec:method:phase_estimation} and ~\ref{sec:method:gating}. We then use Nadarya-Watson kernel regression\cite{Bishop2006} to learn a phase-parametrized image manifold in the form of a function $M(\phi): [0, 1) \to R^m $ that maps any cardiac phase $\phi$ to an image using kernel-weighted local average as follows:
\begin{equation}
M(\phi) = \frac{\sum_{t = 1}^{N} K \left( \phi, \hat{\phi}_{heart}(t) \right) I(t)}{\sum_{t = 1}^{n} K \left( \phi, \hat{\phi}_{heart}(t) \right)} 
\end{equation}
where kernel $K\left( \phi, \hat{\phi}_{heart}(t) \right) = exp\left \{ -\frac{ \mid \phi - \hat{\phi}_{heart}(t) \mid^2}{2  \sigma^2_\phi} \right \} \times exp\left \{ -\frac{ \mid L(\phi) - \hat{r}_{heart}(t) \mid^2}{2  \sigma^2_{L}} \right \}$ is defined as the product of two radial-basis function (RBF) kernels. The first RBF kernel  weighs images inversely proportional to their distance (accounting for periodicity) in the cardiac phase space. Its bandwidth $\sigma_\phi$ is set equal to a constant $k_\phi = 0.4$ times the median difference in cardiac phase between consecutive frames of the given image sequence. The second RBF kernel weighs images inversely proportional to their deviation from the LOWESS prediction $L(\phi)$ in the residual/heart-beat signal space. Its bandwidth $\sigma_{L}$ is set equal to a constant $k_L = 2.0$ times the robust estimate of standard deviation/dispersion $\hat{\sigma}_{L}$ of non-respiratory frames around the LOWESS fit as described in Section~\ref{sec:method:gating}.

	A representative single-cycle video of the subject's heart-beat can now be reconstructed by sampling the manifold $M(\phi)$ at any desired resolution from the range $[0, 1)$ of the cardiac phase. Note that this approach will facilitate temporal super-resolution only if the set of images observed in each distinct period of the given image sequence differ in cardiac phase by some non-zero amount. Also, the amount of temporal super-resolution possible will depend both on the number of cycles captured and the amount by which one can expect the corresponding images to be perturbed in phase across cycles. 
%\vspace{-0.5cm}
\section{Results}
\label{sec:results}
%
%
\begin{IonFigT}
\centering
%
\includegraphics[width=5.0in]{figures/decoded/2015-07-27-10-36-06_2015-07-15-16-56-16_1.raw.bmode/phaseordered.png}
\ionbox{5.0in}\\
%
\caption{Visualization of m-mode frames ordered by estimated cardiac phase}
\label{fig:phase_ordering}
\end{IonFigT}
%
%
%\vspace{-0.5cm}
\section{Discussion and future work}
\label{sec:discussion}
%
